---
title: "OWID Monkey Pox"
author: "David Jackson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(forecast)
library(scales)
library(lubridate)
```

## Our World In Data: Monkey Pox Worldwide (Up to Aug. 04,2022)

```{r}
rm(list=ls())
monkey_pox <- read_csv("https://raw.githubusercontent.com/owid/monkeypox/main/owid-monkeypox-data.csv") %>% janitor::clean_names() %>% filter(date <="2022-08-04")
dim(monkey_pox)
```
```{r}
summary(monkey_pox)
```

### Calculate Total Worldwide Cases and Deaths by Date

#### Cases

```{r}
monkey_pox$wday <- wday(monkey_pox$date)

worldwide_wday <- monkey_pox %>%
    group_by(wday) %>%
    summarise(Cases = sum(new_cases),
              Deaths = sum(new_deaths)) %>% ungroup()

ggplot(worldwide_wday) + geom_col(aes(x=wday, y=Cases))
```


```{r}
worldwide_daily <- monkey_pox %>%
    group_by(date) %>%
    summarise(Cases = sum(new_cases))
```




```{r}
worldwide_daily %>% ggplot() + geom_line(aes(x=date,y=Cases)) +
  labs(title="Worldwide Cases by Date")
```
```{r}
worldwide_daily$Cumcases <- cumsum(worldwide_daily$Cases)
```
```{r}
worldwide_daily %>% ggplot() + geom_line(aes(x=date,y=Cumcases)) +
  labs(title="Worldwide Cumulative Total Cases by Date")
```

#### Deaths

```{r}
monkey_deaths <-   monkey_pox %>% na.omit() %>%
    group_by(date) %>%
    summarise(Deaths = sum(new_deaths))       
monkey_deaths$Cumdeaths <- cumsum(monkey_deaths$Deaths)
```
```{r}
ggplot(monkey_deaths)  + geom_line(aes(x=date,y=Deaths))
ggplot(monkey_deaths)  + geom_line(aes(x=date,y=Cumdeaths))
```

### Begin Models: Worldwide Daily Models


```{r}
Y <- ts(worldwide_daily$Cases,start=c(2022,05,06),frequency = 365)
class(Y) # Class ts
```


```{r}
autoplot(Y) + 
  labs(title ="Worldwide Monkey Pox",y="Daily Cases ")
```
```{r}
DY <- diff(Y)

autoplot(DY) + 
  labs(title ="Wroldwide Monkey Pox: Difference Plot ")
```
```{r}
## Didn't work
# ggseasonplot(DY) +
# labs(title = "Seasonial Plot: Change In Daily Sales")
```
```{r}
## Ddon't work
# ggsubseriesplot(DY)
```

## Forecast with various methods

#### Mean Forecast

```{r}
fit_mean <- meanf(worldwide_daily$Cases)
```

```{r}
checkresiduals(fit_mean)
# forecast_mean <- forecast(fit_mean,h=90)
# autoplot(forecast_mean)
```

#### Fit Seasonal Naive: Didn't work

```{r}
# fit_snaive <- snaive(DY) # Residual sd:
# head(fit_snaive)
```
```{r}
# summary(fit_snaive)
```
```{r}
# checkresiduals(fit_snaive)
# autoplot(fit_snaive)
```

### ETS forecast

```{r}
fit_ets <- ets(Y) # sigma:  0.1077
summary(fit_ets)
```
```{r}
checkresiduals(fit_ets)
```
```{r}
autoplot(fit_ets)
```
```{r}
fcat_ets <- forecast(fit_ets,h=60)
```
```{r}
autoplot(fcat_ets)
```
```{r}
autoplot(fcat_ets,include=36) + labs(title ="4 Years")
```

## Forecast Cumulative Worldwide Cases


```{r}
YY <- ts(worldwide_daily$Cumcases,start=c(2022,05,06),frequency = 365)
class(YY) # Class ts
```

```{r}
autoplot(YY) + 
  labs(title ="Worldwide Monkey Pox",y="Cumulative Cases ")
```

```{r}
DY <- diff(YY)

autoplot(DY) + 
  labs(title ="Wroldwide Monkey Pox: Difference Plot ")
```

## Forecast with various methods

#### Mean Forecast

```{r}
fit_mean <- meanf(worldwide_daily$Cumcases)
head(fit_mean)
```
```{r}
summary(fit_mean)
```
```{r}
checkresiduals(fit_mean)

# forecast_mean <- forecast(fit_mean,h=90)
# autoplot(forecast_mean)
```

#### Fit Seasonal Naive: Didn't work

```{r}
# fit_snaive <- snaive(DY) # Residual sd:
# head(fit_snaive)
```
```{r}
# summary(fit_snaive)
```
```{r}
# checkresiduals(fit_snaive)
```

### ETS forecast

```{r}
fit_ets <- ets(YY) # sigma:  0.1077
summary(fit_ets)
```
```{r}
checkresiduals(fit_ets)
```
```{r}
autoplot(fit_ets)
```
```{r}
fcat_ets <- forecast(fit_ets,h=30)
```
```{r}
autoplot(fcat_ets) +
  scale_y_continuous(label = comma) +
  labs(title ="Wordwide Monkey Pox Forecast: Aug. 05,2022",
       subtitle = "(100,000 by Sept. 5, 2022)")
```

### Facebook Prophet

```{r}
library(prophet)
df <- monkey_pox %>% select(date,total_cases) %>%
  rename(ds =date,y=total_cases)
```
```{r}
m <- prophet(df, weekly.seasonality = 5)
future <- make_future_dataframe(m, periods = 30)
tail(future)
```
```{r}
forecast <- predict(m, future)
tail(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')])

```
```{r}
plot(m, forecast)

```

```{r}
prophet_plot_components(m, forecast)
```
```{r}
dyplot.prophet(m, forecast)
```

