---
title: 'Forecast: Super Store Sales'
author: "David Jackson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp2) # Includes forecast package
library(tidyverse)
library(DBI)
library(odbc)
library(janitor)
# library(fredr)
library(lubridate)
```

## Pull SuperStore Sales date

```{r}
rm(list = ls())
con <- DBI::dbConnect(odbc::odbc(), 
                      Driver = "SQL Server", 
                      Server = "localhost\\SQLEXPRESS", 
                      Database = "superstore", 
                      Trusted_Connection = "True")

```

```{r}

orders <- dbGetQuery(con,"select * from orders")
orders <- janitor::clean_names(orders)
```

## Calculate Monthly Sales

```{r}
monthly <- orders %>% 
  group_by(month =floor_date(order_date,"month")) %>%
    summarise(Sales = sum(sales))

```

```{r}
summary(monthly)
```

## Quick plot of sales by day

```{r}
ggplot(monthly) + geom_line(aes(x=month,y=Sales)) +
  labs(title ="Sales ($) By Month",y="Sales ($)")
```

## Begin Forecast

### Create Time Series

```{r}
Y <- ts(monthly$Sales,start=c(2011,1),frequency = 12)
class(Y) # Class ts
```

```{r}
autoplot(Y) + 
  labs(title ="Time Plot: Super Store Monthly Sales",y="Millions of USD ")
```

### Data has as strong trend. Calculate Differnce

```{r}
DY <- diff(Y)

autoplot(DY) + 
  labs(title ="Time Plot: Change in US Retail Sales ")
```

### Data Appears trend stationary: Use to investigate seasonality.

```{r}
ggseasonplot(DY) +
  labs(title = "Seasonial Plot: Change In Daily Sales")
```

### Seasonial Plot: sub-series

```{r}
ggsubseriesplot(DY)
```

## Forecast with various methods
### Seasonal Naive (snaive)

```{r}
fit <- snaive(DY) # Residual sd:
head(fit)
```
```{r}
summary(fit)
```
```{r}
checkresiduals(fit)
autoplot(fit)
```

### ETS forecast

```{r}
fit_ets <- ets(Y) # sigma:  0.1077
summary(fit_ets)
```
```{r}
checkresiduals(fit_ets)
```
```{r}
autoplot(fit_ets)
```
```{r}
fcat_ets <- forecast(fit_ets,h=12)
```
```{r}
autoplot(fcat_ets)
```
```{r}
autoplot(fcat_ets,include=12) + labs(title ="4 Years")
```

## Begin Auto Arima Model

### Create Fit for Model

```{r}
fit_arima <- auto.arima(Y,d=1,D=1,stepwise = FALSE,approximation = FALSE, trace= TRUE)
```

```{r}
summary(fit_arima)
```
```{r}
checkresiduals(fit_arima)
```

### Forecast and Plot Arima Model

```{r}
fcast1 <- forecast(fit_arima,h=24)
```
```{r}
autoplot(fcast1)
```
```{r}
autoplot(fcast1,include=24) + labs(title = "Last 3 Years") 
```
