---
title: "OWID USA Monkey Pox"
author: "David Jackson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(forecast)
library(scales)
```

## Our World In Data: Monkey Pox United States

```{r}
rm(list=ls())
monkey_pox <- read_csv("https://raw.githubusercontent.com/owid/notebooks/main/EdouardMathieu/monkeypox/owid-monkeypox-data.csv") %>% janitor::clean_names() %>% filter(location =="United States")
dim(monkey_pox)
```
```{r}
summary(monkey_pox)
```

### Calculate Total United States Cases and Deaths by Date

#### Cases

```{r}
monkey_daily <- monkey_pox %>%
    group_by(date) %>%
    summarise(Cases = sum(new_cases),
              Deaths = sum(new_deaths)) %>% ungroup()
```

```{r}
monkey_daily %>% ggplot() + geom_line(aes(x=date,y=Cases)) +
  labs(title="US Cases by Date")
```
```{r}
monkey_daily$Cumcases <- cumsum(monkey_daily$Cases)
```
```{r}
monkey_daily %>% ggplot() + geom_line(aes(x=date,y=Cumcases)) +
  labs(title="US Cumulative Total Cases by Date")
```

#### Deaths

```{r}
monkey_daily %>% ggplot() + geom_line(aes(x=date,y=Deaths)) +
  labs(title="United States Deaths by Date")
```
```{r}
monkey_daily$Cumdeaths <- cumsum(monkey_daily$Deaths)
```
```{r}
monkey_daily %>% ggplot() + geom_line(aes(x=date,y=Cumdeaths)) +
  labs(title="United States Cumulative Total Deaths by Date")
```

### Begin Models: United States Daily Models


```{r}
Y <- ts(monkey_daily$Cases,start=c(2022,05,06),frequency = 365)
class(Y) # Class ts
```


```{r}
autoplot(Y) + 
  labs(title ="United States Monkey Pox",y="Daily Cases ")
```
```{r}
DY <- diff(Y)

autoplot(DY) + 
  labs(title ="Wroldwide Monkey Pox: Difference Plot ")
```
```{r}
## Didn't work
ggseasonplot(DY) +
  labs(title = "Seasonial Plot: Change In Daily Sales")
```
```{r}
## Ddon't work
ggsubseriesplot(DY)
```

## Forecast with various methods

#### Mean Forecast

```{r}
fit_mean <- meanf(monkey_daily$Cases)
head(fit_mean)
```
```{r}
summary(fit_mean)
```
```{r}
checkresiduals(fit_mean)
# forecast_mean <- forecast(fit_mean,h=90)
# autoplot(forecast_mean)
```

#### Fit Seasonal Naive: Didn't work

```{r}
# fit_snaive <- snaive(DY) # Residual sd:
# head(fit_snaive)
```
```{r}
# summary(fit_snaive)
```
```{r}
# checkresiduals(fit_snaive)
# autoplot(fit_snaive)
```

### ETS forecast

```{r}
fit_ets <- ets(Y) # sigma:  0.1077
summary(fit_ets)
```
```{r}
checkresiduals(fit_ets)
```
```{r}
autoplot(fit_ets)
```
```{r}
fcat_ets <- forecast(fit_ets,h=30)
```
```{r}
autoplot(fcat_ets)
```
`

## Forecast Cumulative United States Cases


```{r}
YY <- ts(monkey_daily$Cumcases,start=c(2022,05,06),frequency = 365)
class(YY) # Class ts
```

```{r}
autoplot(YY) + 
  labs(title ="United States Monkey Pox",y="Cumulative Cases ")
```

```{r}
DY <- diff(YY)

autoplot(DY) + 
  labs(title ="US Monkey Pox: Difference Plot ")
```

## Forecast with various methods

#### Mean Forecast

```{r}
fit_mean <- meanf(monkey_daily$Cumcases)
head(fit_mean)
```
```{r}
summary(fit_mean)
```
```{r}
checkresiduals(fit_mean)

# forecast_mean <- forecast(fit_mean,h=90)
# autoplot(forecast_mean)
```

#### Fit Seasonal Naive: Didn't work

```{r}
# fit_snaive <- snaive(DY) # Residual sd:
# head(fit_snaive)
```
```{r}
# summary(fit_snaive)
```
```{r}
# checkresiduals(fit_snaive)
```

### ETS forecast

```{r}
fit_ets <- ets(YY) # sigma:  0.1077
summary(fit_ets)
```
```{r}
checkresiduals(fit_ets)
```
```{r}
autoplot(fit_ets)
```
```{r}
fcat_ets <- forecast(fit_ets,h=14)
```
```{r}
autoplot(fcat_ets)
```
```{r}
autoplot(fcat_ets,include=36) + labs(title ="4 Years")
```
